---
title: "P8105 Homework 2"
author: Adam Whalen
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trash Wheel dataset.

```{r trashwheel_import}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
    )
```

Read precipitation data for 2018 and 2017

```{r precip_import}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation.

```{r precip_merge}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trash Wheel trash collector in Baltimore, MD. As trash flows to the Inner Harbor, Mr. Trash Wheel collects that trash and stores it in a dumpster. This dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include monthly precipitation data. 

## Problem 2

Read and clean the NYC Transit data set.

```{r transit_import}
transit_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>%
  mutate(
    entry = recode(entry, "YES" = 1, "NO" = 0),
    entry = as.logical(entry)
  )
```

The NYC Transit data set contains information about all entry and exit locations for every subway station in New York. The data set consists of `r nrow(transit_df)` observations and `r ncol(transit_df)` variables, including location, line, routes served, and ADA compliance to name a few. To clean the dataset, we tidied variable names, retained only select important variables, and converted a character variable to logical. However, these data are not completely tidy yet, as we have multiple columns that pertain to the same piece of information (route), for example. 

#### Some answers to some questions:

1) There are `r count(distinct(transit_df, station_name, line))` distinct stations in this data set.
2) `r nrow(distinct(filter(transit_df, ada == "TRUE"), station_name, line))` stations are ADA compliant.
3) `r format(round((nrow(filter(transit_df, entry == "TRUE" & vending == "NO")) / nrow(filter(transit_df, vending == "NO"))) * 100, 2), nsmall = 2)` percent of entrances and exits that do not have ticket vending allow entrance.

Now, reformat the data set by route name variables.
```{r transit_tidying}
transit_tidy =
  transit_df %>% 
  mutate(
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11),
    ) %>% 
  pivot_longer(
    cols = route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  ) %>% 
  drop_na(route_name)
```

There are `r nrow(distinct(filter(transit_tidy, route_name == "A"), station_name, line))` distinct stations that serve the A line. Of those that serve the A, `r nrow(distinct(filter(transit_tidy, route_name == "A" & ada == "TRUE"), station_name, line))` stations are ADA compliant.

## Problem 3

First, import and clean the `pols` data set.

```{r pols_import_tidy}
pols_df = 
  read_csv("./data/pols-month.csv") %>% 
  separate(mon, c("year", "month", "day")) %>% 
  mutate(
    month = as.integer(month),
    year = as.integer(year),
    day = as.integer(day),
    president = case_when(
      prez_gop == 1 ~ "gop",
      prez_dem == 1 ~ "dem"
    )
  )

pols_tidy = 
  left_join(pols_df, month_df, by = "month") %>% 
    relocate(year, month_name) %>% 
    select(-c(month, day, prez_gop, prez_dem))
```

Next, import and clean the `snp` data set.
```{r snp_import_tidy}
snp_df = 
  read_csv("./data/snp.csv") %>% 
  separate (date, c("month", "day", "year"), sep = "/") %>% 
  mutate(
    month = as.integer(month),
    year = as.integer(year),
    day = as.integer(day)
  )

snp_tidy = 
  left_join(snp_df, month_df, by = "month") %>% 
  relocate(year, month_name) %>% 
  select(-c(day, month))
```

Then, import and clean the `unemployment` data set.
```{r unemp_import_tidy}
unemployment_df = 
  read_csv("./data/unemployment.csv") %>% 
  pivot_longer(
    cols = Jan:Dec,
    names_to = "month_abb",
    values_to = "unemp_rate"
  ) %>% 
  janitor::clean_names()

month_df = 
  month_df %>% 
    mutate(month_abb = month.abb)

unemployment_tidy = 
  left_join(unemployment_df, month_df, by = "month_abb") %>% 
  relocate(year, month_name) %>% 
  select(-c(month_abb, month))
```

Finally, merge the `pol` and `snp` data sets, then merge `unemployment`.
```{r complete_merge}
complete_df = 
  left_join(pols_tidy, snp_tidy, by = c("year", "month_name")) %>% 
  left_join(unemployment_tidy, by = c("year", "month_name"))
```

These data sets come from the political statistics company FiveThirtyEight and consist of various measures related to government and politics in the United States. The `pols` data set contains information about national elected officials and their party affiliations since 1947. The `snp` data set contains information about closing market value for the S&P 500 by year and month. The `unemployment` data set contains information about the national unemployment rate in the US by month and year. Our combined data set has `r nrow(complete_df)` monthly observations and `r ncol(complete_df)` variables, compiled from the three individual sets. Important variables include president's political party, unemployment rate, and close value for the S&P.
